---
title: "HWK2"
author: "Felipe Pena Mosca"
date: "2/3/2022"
output: html_document
---

```{r setup, include=FALSE}
#Load packages

library(tidyverse)

#Working directory for .RMD

knitr::opts_knit$set(echo = TRUE,
                     root.dir = rprojroot::find_rstudio_root_file())

#Controlling figure output in markdown

knitr::opts_chunk$set(
#  fig.height =   
  fig.width = 6,
#  fig.asp = .5,
  out.width = "90%",
#  out.height = 
 fig.align = "center",
  cache = FALSE,
  echo  = TRUE
)

#Set Theme for ggplot2 - centers title and legend at bottom by default

theme_classic()

theme_set(theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom"))

#Set Scientific notation output and decimal places for knitr

options(scipen = 999)
options(digits = 4)

```

Part 1

```{r Part 1, include=TRUE}
# x = a random sample of size n = 1000 from N(0, 1) (rnorm())

x <- rnorm(1000,0,1)

# y = a random sample of size n = 1000 from N(1, 2)

y <- rnorm(1000,1,2)

# Create vector that combines x+y

z <- y + x

# sum_indicator = a logical variable, TRUE if x + y > 0.5 (hint: ifelse())

sum_indicator <- ifelse(z > 0.5,"Yes","No")

# Generate dataframe (need to use tibble for this)

test <- data.frame(x,y,sum_indicator)

# Relevel variable

test$sum_indicator <- forcats::fct_relevel(test$sum_indicator,"Yes","No")

# Scatter plot y by x

ggplot(test,aes(y = y,x = x,color = sum_indicator)) + 
  geom_point() +
  labs(title="Scatterplot of x by y",x ="Random Sample N(0, 1)", y = "Random Sample N(1, 2)",colour="x+y > 0.5")

```

Part 2.1

This dataset includes 344 rows and 17 variables.

Variables included:

- Sample: Number an integer denoting the continuous numbering sequence for each sample.

- Species: a character string denoting the penguin species
Region a character string denoting the region of Palmer LTER sampling grid.

- Island: a character string denoting the island near Palmer Station where samples were collected.

- Stage: a character string denoting reproductive stage at sampling.

- Individual-ID: a character string denoting the unique ID for each individual in dataset.

- Clutch Completion: a character string denoting if the study nest observed with a full clutch, i.e., 2
eggs.

- Date Egg: a date denoting the date study nest observed with 1 egg (sampled).

- Culmen Length: a number denoting the length of the dorsal ridge of a bird’s bill (millimeters).

- Culmen Depth: a number denoting the depth of the dorsal ridge of a bird’s bill (millimeters).

- Flipper Length: an integer denoting the length penguin flipper (millimeters)
Body Mass an integer denoting the penguin body mass (grams).

- Sex: a character string denoting the sex of an animal
Delta 15 N a number denoting the measure of the ratio of stable isotopes 15N:14N.

- Delta 13 C: a number denoting the measure of the ratio of stable isotopes 13C:12C.

- Comments: a character string with text providing additional relevant information for data

```{r Part 2.1, include=TRUE}
# Part 2.1

# Upload data

penguin.df <- read_rds("./data/penguin.RDS")

#The number of observations (rows), variables/features (columns), and what each variable/feature describes about each observation

# Mean and standard deviation flipper length

mean(penguin.df$flipper_length_mm,na.rm = TRUE)
sd(penguin.df$flipper_length_mm,na.rm = TRUE)

# Mean and standard deviation bill length

mean(penguin.df$bill_length_mm,na.rm = TRUE)
sd(penguin.df$flipper_length_mm,na.rm = TRUE)

```

Part 2.2

```{r Part 2.2, include=TRUE,warning = FALSE}

# Part 2.2

# Scatter plot bill length by flipper length and species

ggplot(penguin.df,aes(y =flipper_length_mm,x =bill_length_mm,color=species)) + geom_point()+
labs(colour="Species")+labs(title="Scatterplot of flipper length by bill length and penguin species",x ="Bill length (mm)", y = "Flipper length (mm)")

# Scatter plot bill length by flipper length and sex

ggplot(subset(penguin.df,!is.na(sex)),aes(y =flipper_length_mm,x =bill_length_mm,color=sex)) + geom_point() + labs(colour="Sex")+labs(title="Scatterplot of flipper length by bill length and penguin sex",x ="Bill length (mm)", y = "Flipper length (mm)")+ 
geom_smooth(formula = y ~ x, 
method  = "lm", 
se      = FALSE)+
facet_grid(~sex)

```
This scatterplot shows a positive association between bill length (mm) (independent variable) and flipper length (mm) (dependent variable) in both female and male penguins

